Requirements and Current State of Knowledge Representation Systems for Service Robotics

### 2012-05-11 krs-survey.tex
Refined and completed section on interaction

In the survey inclusion criteria, we have insisted on only considering robotic
systems that acquire knowledge by themselves, during their runtime.

\emph{Acquiring knowledge} means in our context building new statements
(usually as new logical facts) connected to our existing knowledge from
external sources of information. We consider mainly three of them:
proprioceptive/exteroceptive sensors, interaction with other agents, humans or
robots, and remote databases. This process has generally at least two steps:
the information acquisition by itself, and the \emph{transformation} of the
information into knowledge, \emph{aligned} with the robot existing model
(following our terminology for information and knowledge, as discussed in the
introduction).

Knowledge acquisition is generally not done directly in the knowledge
representation system. On the contrary, several, if not numerous, external
components are usually required to convert percepts into symbolic facts and to
ground them.

@@

@@

@@

\subsubsection{Language Processing}
\label{sect|nlp}

The scope of this survey is service robotic, which suggest human-robot
interaction. While numerous modalities of communication between humans and
robots do exist and have been extensively studied, the most challenging of
these interactions is natural, bidirectional, unconstrained verbal
communication. Kruijff et al. provides in~\cite{Kruijff2010} an up-to-date
survey of literature on situated human-robot dialogue, focusing on formal
representation systems, bi-directionality of the interaction and context
building.

>>>>

In the survey inclusion criteria, we have chosen to only consider robotic
systems that acquire knowledge by themselves, at runtime.

In our context, \emph{acquiring knowledge} means to build new logical
statements from data sources and to anchor them into the existing knowledge. We
consider three possible sources of data: proprioceptive/exteroceptive sensing,
interaction with other agents, humans or robots, and remote knowledge bases.
The acquisition process has generally two steps: the information acquisition by
itself, and the \emph{transformation} of the information into knowledge,
\emph{aligned} with the robot existing model (following our terminology for
information and knowledge, as discussed in the introduction).

It must be observed that knowledge acquisition is generally not done directly
in the knowledge representation system. External components (often aggregated
into knowledge acquisition \emph{pipelines}) are usually required to convert
percepts into symbolic facts and to ground them.

@@

From the point of view of knowledge representation, the sensing capability can
be split into proprioceptive sensing (\ie, sensing of the robot own internal
state) and exteroceptive sensing (sensing of the robot environment). The
(physical) introspection capabilities of the robot relies on the former.

\fxwarning{What to say here?}

@@

Interaction with other intelligent agents (humans or robots) is another
important source of knowledge acquisition. It relies obviously on some form of
sensing (from speech recognition to gesture recognition) but we distinguish it
from the previous section because \emph{interaction} implies a form of
communication. Communication is associated to specific functions
(figure~\ref{fig|communication-model}), and in particular, it implies a shared
\emph{context} (usually implicit) between interactors.

\begin{figure}
    \centering
    \includegraphics[width=0.8\columnwidth]{jakobson_communication_model.pdf}

    \caption{The Jakobson~\ref{Jakobson1960} communication model. In italics,
    the associated \emph{communication functions}.}

    \label{fig|communication-model}
\end{figure}

We distinguish between two main interaction channels: verbal and deictic.

\begin{scriptsize}
\begin{center}
\begin{tikzpicture}[taxonomy]
    \node [taxon] {\bf D.1.2 Interaction}
            child {node [taxon] {{\bf D.1.2.2} Deictic Interaction}}
            child {node [taxon] {{\bf D.1.2.1} Verbal Interaction}};
\end{tikzpicture}
\end{center}
\end{scriptsize}

\label{sect|nlp}

The field of verbal interaction processing for robots spans from pattern-based,
constrainted sentences recognition to natural, bidirectional, unconstrained
verbal communication. Kruijff et al. provides in~\cite{Kruijff2010} an
up-to-date survey of literature on situated human-robot dialogue, focusing on
formal representation systems, bi-directionality of the interaction and context
building.

Deictic (used in the literal meaning of ``display, demonstration, reference'')
interaction is also an established field of research in human-robot
interaction. Common deictic forms of communication include attentional focus
(via face and gaze tracking) and joint attention, pointing, emotional
expressions (based on face expressions, postures, emotional gestures).

Like other knowledge acquisition modalities, the recognition and interpretation
of deictic communication is rarely directly included in a KRS, but, as
previously mentioned, the symbolic representation of such communication acts is
relevant and important to achieve successful multi-modal interactions.

@@
